%% Title
\titlepage[of The University of Sydney]{%
  A dissertation submitted for partial fulfillment of the requirements \\ for the degree of Doctor of Philosophy}

%% Abstract
\begin{abstract}%[\smaller \thetitle\\ \vspace*{1cm} \smaller {\theauthor}]
After the discovery of the Higgs boson at LHC \cite{RN62,RN63} and the subsequent measurements of its mass and couplings, the search for new physics has become more important than ever, with one of the most promising candidates being theories which exhibit supersymmetry (\acrshort{susy}). A space-time symmetry between fundamental integer and half-integer spin particles, SUSY proposes a plethora of new states which may be in reach of present-day collider technology, with even greater prospects for those of the future. As a consistent solution to the hierarchy problem, the unnaturally large quantum corrections to the bare Higgs boson mass from short-distance scale physics, this thesis explores anomalies from experiment left unexplained by the enormously successful Standard Model of particle physics, offered consistently even in minimal supersymmetric extensions.

With this in mind, we initially focus on the predictions from the Minimal Supersymmetric Standard Model (\acrshort{mssm}) without assumption of specific SUSY breaking mechanisms. Our MSSM phenomenology explores two of the most sensitive observables - the muon anomalous magnetic moment and dark matter. In conjunction with collider and other searches, we find a number of parameter regions still viable, though future 100 TeV collider searches may even be able to close the case on the MSSM explanation for these anomalies. Subsequently, we explore the idea of `naturalness' (or fine-tuning) in the MSSM. In light of current experimental limits on coloured sparticles, we believe naturalness considerations strongly hint the effectiveness of the MSSM up to new physics scales as low as 100 TeV.

Finally, we propose simple modification to minimal supersymmetry and its place in the early universe, without augmenting the gauge structure or particle content, in order to alleviate constraint on the allowable parameter space of the MSSM. The former focuses on the more comfortable accommodation of a 125 GeV Higgs boson mass within the framework of the MSSM through non-linear realization of electroweak gauge symmetry, whilst the latter accounts for a temporary cosmological dark matter (\acrshort{dm}) decay phase, avoiding the commonly encountered overabundance of DM in MSSM parameter regions.

Our results seem to yearn for higher center-of-mass energy collisions and more precise experimental observations in order to continue setting our targets and close in on supersymmetric models.
\end{abstract}

%% Declaration
\begin{declaration}
\noindent
This thesis presents new findings for the Minimal Supersymmetric Standard Model (\acrshort{mssm}), in a general model-independent fashion using a combination of analytical and numerical methods. It is the final piece of work towards completion of my doctorate in Theoretical Particle Physics at the University of Sydney.

Chapter \ref{chap:muong-2} is based on the published work in \cite{RN2} with corresponding conference notes in \cite{RN760}. I performed all major calculations for the low-energy observables, dark matter properties, and collider simulations and produced plots for the manuscript.

Chapter \ref{chap:finetuning} is based on the published work in \cite{RN803}. I performed all calculations and produced the plots as well as some contribution to the text in the manuscript.

Chapter \ref{chap:wimpdecay} is based on the published work in \cite{RN802}. I calculated some numerical results, derivations and plots.

Chapter \ref{chap:nonlinearhiggs} is based on the published work in \cite{RN1}. I calculated the potential contributions and mass spectrum for the manuscript.

Other research output during this doctorate that will not be fully discussed in this thesis are published in \cite{RN763,RN764,RN765}.
\newpage
\hspace{0pt}
\vfill
%\noindent
\textit{This certifies that the content of this thesis is entirely my own work, except where explicit reference is made to the work of others. This thesis has not been submitted as part of any other degree or qualification for this or any other university. I have acknowledged all sources of help in this thesis.}
  \vspace*{1cm}
  \begin{flushright}
  	\textit{Candidate's signature:}\hspace{0.5cm} \makebox[1.5in]{\hrulefill} \\
    \textit{Matthew Talia}\hspace{0.5cm}\, \\
    \vspace*{0.75cm}
    \textit{Supervisor's signature:}\hspace{0.5cm} \makebox[1.5in]{\hrulefill} \\
    \textit{Archil Kobakhidze}\hspace{0.3cm}\,
  \end{flushright}
\vfill
\hspace{0pt}
\end{declaration}

%% Acknowledgements
\begin{acknowledgements}
\noindent
Apologies in advance for this exhaustive, but necessary list. This thesis, amongst some other phenomena throughout time like neutron star mergers and actor Bill Paxton, is of course only made possible by all the fundamental particles and their interactions.

It is with immense gratitude that I must first acknowledge my advisor, Archil Kobakhidze for his support in all my endeavours without hesitation. With your guidance, I have learned to be humble in my efforts and to approach all science with a healthy dose of critical thought and skepticism. On a side note, Archil also promised that my dissertation would be accepted provided that I turned up to our regular soccer matches\footnote{Proper written citation for this statement pending.}.

I would also like to thank Michael Schmidt for his co-supervision and interesting discussions at many points during my doctoral studies. Special thanks to Lei Wu for help in developing my early interests in collider studies and computational work. And on that note, to Goncalo Borges for putting up with my continued (and at times, intensive) use of the computing cluster.

Thanks to those who supported me in my travels around Spain, northern Italy and the UK to present my work, particularly Zurab Berezhiani, Dmitri Sorokin and Francesco Coradeschi for invited talks. I should also mention the ARC Center of Excellence for Particle Physics at the Terascale (CoEPP) for supporting this research from the beginning.

It has been a privilege to work with some excellent minds in Room 342 over the years, and I could see myself benefiting greatly from some fantastic theoretical discussions, especially Neil Barrie, Sundy Arunasalam, Cyril Lagger and anyone else I have forgotten. Most of all, during my time in Sydney, I got to make some great friends (Lachlan and Abed) whom I will never forget. And of course, everyone back home in Melbourne, some of whom I've known almost all my life.

To Mum and Dad and all the family back home, thank you for all the support from such a distance. It is because of you that I made it here, safe and sound. To Dean, you made it to the finish line first, I just kind-of followed along hoping that I had the potential to do the same. And of course, to my partner-in-crime Sophie who was there every (and I mean every) step of the way. Fine, I'll say thanks to our cat, Newton. Is that better now?

To close this especially important section, it is often a thought of mine that we seldom remember to thank ourselves for overcoming our own trials and tribulations. Along with the special help of friends, family, my partner, members of the department, and other important people along the way, I was able to slowly navigate the treacherous dark tunnel that is mental health recovery and emerge from the other side - something I thought I would never be capable of.

\end{acknowledgements}

% Preface
%\begin{preface}
%
%\end{preface}

%%Acronyms
\clearpage
 
\printglossary[type=\acronymtype,nonumberlist]

%\printglossary[type=\acronymtype]
 
%\printglossary

% Given a set of numbers, there are elementary methods to compute 
% its \acrlong{gcd}, which is abbreviated \acrshort{gcd}. This 
% process is similar to that used for the \acrfull{lcm}.

%% ToC
\tableofcontents

%% Strictly optional!
\frontquote{ \centering
I'm not afraid of death. I'm an old physicist - I'm afraid of time.
 }%
  {Dr. Brand, from \textit{Interstellar (2014) \cite{RN759}.}}
%% I don't want a page number on the following blank page either.
\thispagestyle{empty}

%%Outline
\input{Outline}